---
title: "alessandra_puig-santana_snowshoe_report"
author: "Alessandra Puig-Santana"
date: "11/7/2021"
output: html_document
code_folding: hide 
---
### Overview 
This reports explores the juvenile snowshoe hare population at Bonanza Creek from 1999 - 2012.

### Citation
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14 

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(ggbeeswarm)
library(kableExtra)
library(effsize)
library(broom)
```

```{r}
# Read in data
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv")) %>%
  mutate(sex = case_when( # Changed names in sex observations 
    sex %in% "f" ~ "Female",
    sex %in% "m" ~ "Male"))
```

## Juvenile Hare Initial Investigation
Annual juvenile hare trap counts. Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 

### Data and Analysis
```{r}
j_hares <- bonanza_hares %>%
  filter(age == "j") %>% # Filter for juveniles in the age column
  mutate(date = mdy(date)) %>% # Edited date column to have date in yyyy-dd-mm
  mutate(year = year(date)) # Created new column with the year 

j_hares_count <- j_hares %>% #Subset of data counting the amount of hares per year
  count(year)

# Annual juvenile hare trap counts 
j_hares_summary <- j_hares_count %>% # Summary table statistics including mean, median, max, and min
  summarize(j_hares_mean = mean(n, na.rm = TRUE),
            j_hares_median = median(n, na.rm = TRUE),
            j_hares_max = max(n, na.rm = TRUE),
            j_hares_min = min(n, na.rm = TRUE))

# Creating summary table 
j_hares_summary_table <- j_hares_summary %>%
  kbl() %>%
  kable_styling(full_width = FALSE, font = 13)
```

```{r, fig.cap= "Figure 1: This bar graph demonstates the amount of juveniles snowshoe hares collected from 1999 to 2012."}
ggplot(data = j_hares_count, aes(x = year, 
                           y = n)) +
  geom_line(color = "darkgreen") +
  scale_x_discrete(limit = c(1999:2012)) +
  labs(x = "Year",
       y = "Count",
       title = "Amount of Juvenile Snowshoe Hares collected from 1999-2012") +
  theme_minimal()
```

The average of juvenile hares trapped through 1999 to 2012 `r round(j_hares_summary$j_hares_mean, 0)`. The median of juvenile hares `r round(j_hares_summary$j_hares_median, 0)`. The maximum amount of hares caught were in 1999 with `r j_hares_summary$j_hares_max` and a minimum in 2010 with `r j_hares_summary$j_hares_min`. 

## Visualize juvenile hare weights
Create a finalized data visualization (or visualizations) in which you compare **juvenile hare weights by sex and site (grid)** (ignore other variables). You should include groups where these variables are not recorded (`NA`). Include finalized figure captions below your visualization(s).

Then, in 2 - 3 sentences below your visualization(s) describe in text the main takeaways that you would want the audience to understand from your figures.
```{r, fig.cap = "Figure 2: This scatterplot demonstrates the differences in average weights (g) between male and female juvenile snowshoe hares in the three collection sites."}
sub_juv_hares <- j_hares %>%
  group_by(sex, grid) %>% 
  mutate(grid = case_when(
    grid == "bonbs" ~ "Bonanza Black Spruce",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian"))

# Data Visualization
ggplot(data = sub_juv_hares, aes(x = sex,
                                 y = weight)) +
  geom_jitter(aes(color = sex)) +
  scale_color_manual(values = c("green", "blue"), name = "Sex") +
  geom_boxplot(fill = NA, width = 0.3, outlier.color = NA) +
  stat_summary(fun = mean,
               geom = "point",
               shape = 20,
               size = 4,
               color = "black",
               fill = "black") +
  facet_wrap(~ grid) +
  theme_minimal() +
  labs(title = "Weight differences between male and female snowshoe hares in the three different collection sites")
```

## Juvenile weight comparisons (male and female)
Juvenile weight comparison (male & female snowshoe hares). Further compare mean weights for **juvenile male and female snowshoe hares** included in this study. In this section, you should include:

A finalized table containing the mean, standard deviation, and sample size for male and female juvenile snowshoe hares, with a table caption (remember, a table caption should be placed *above* the table in your report)
```{r}
# Summary statistics
j_hares_sex <- j_hares %>%
  select("sex", "weight") %>%
  group_by(sex) %>%
  drop_na(weight) %>%
  summarize(mean = round(mean(weight), 2),
            sd = round(sd(weight), 2),
            sample = n())

# Summary Statistics table 
j_hares_sex %>% 
  kable(col.name = c("Sex", "Mean (g)", "Standard Deviation (g)", "Sample Size"))
```

### Exploratory data

```{r}
# Statistical data

# Pulling weight vectors for both male and female juvenile snowshoe hares
m_hares <- j_hares %>%
  filter(sex == "Male") %>% 
  drop_na(weight) %>%
  pull(weight)

f_hares <- j_hares %>%
  filter(sex == "Female") %>%
  drop_na(weight) %>%
  pull(weight)

# Plotting exploratory visualizations to check assumptions
# For males: 
hist(m_hares) # Demonstrating normal distribution - looks relatively normal?
qqnorm(m_hares) # Demonstrating linearity - pretty linear?

#For females: 
hist(f_hares) # Negatively skewed
qqnorm(f_hares) # Pretty linear?

# Overall: assumptions of normality hold, & sample size additionally >> Central Limit Theorem (means will be normally distributed). Two-sample t-test OK for means comparison! 

# Running a ttest on male and female juvenile snowshoe hares
j_hares_ttest <- t.test(m_hares, f_hares)

# Tidy model for output on ttest values 
tidy_j_hares_tttest <- tidy(j_hares_ttest)

# Mean and Standard deviations for male and female snowshoe hares
mean_m_j_hares <- mean(m_hares)
sd_m_j_hares <- sd(m_hares)
mean_f_j_hares <- mean(f_hares)
sd_f_j_hares <- sd(f_hares)

# Effect size 
j_hares_effsize <- cohen.d(m_hares, f_hares)
```
** Add analysis ** 

## Relationship between juvenile weight & hind foot length
Relationship between juvenile weight & hind foot length. Explore the relationship between juvenile snowshoe hare hind foot length (variable `hindft`, recorded in millimeters), and weight. Include all juvenile hares in the study. Create a final visualization of juvenile hare hind foot length versus weight. Include a figure caption.

```{r}
# Do linear regression model 
linear_model_j_hares <- lm(weight ~ hindft, data = j_hares)

# Tidy version of model
tidy_linear_model_j_hares <- tidy(linear_model_j_hares)
glance_linear_model_j_hares <- glance(linear_model_j_hares)

# Visualization of linear model
plot(linear_model_j_hares)

# Calculate Pearson's r correlation
cor_j_hares <- cor.test(j_hares$weight, j_hares$hindft)

# Tidy Pearson's r correlation
tidy_cor_j_hares <- tidy(cor_j_hares)
```
```{r}
# Data visualization of model
ggplot(data = j_hares, aes(x = hindft, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(x = "Hind Foot (mm)",
       y = "Weight (g)",
       title = "HFL and Weight")
```

## Summary 
Summary: Include a very short summary describing your main findings from your juvenile hares exploration (bullet points are fine)

## Citations 
Citations: Include citations (professionally formatted), including the data citation and any other resources that were references in your mini report.